---
title: "p8105_hw1_cl4776"
author: "Chenxi Liu"
date: "2025-09-20"
output: github_document
---

# Problem 1

## Install moderndive packages.

## Load dataset
```{r}
library(moderndive)
data("early_january_weather")
```

## Short description using R code.

The `early_january_weather` dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.  
- **time_hour**: date and hour of the observation  
- **temp** (°F): hourly temperature readings  
- **humid** (%): relative humidity  
- **wind_speed**, **wind_dir**, **precip**, **pressure**: additional weather information  
The mean temperature across all observations is `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F.


## Scatterplot of temp vs time_hour, colored by humid.



```{r, message = FALSE}
library(tidyverse)
p_weather = ggplot(
  early_january_weather,
  aes(x = time_hour, y = temp, color = humid)
) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Temperature vs Time (colored by Humidity)",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  ) +
  theme_minimal()

p_weather
```
